SAT Competition 2020\tnoteref{title}

### 2021-01-27 paper/main.tex
track description

\todo{MH?}

@@

\todo{MH?}

@@

\subsection{Similarity of Solvers}
To investigate how similar the solvers of the Main track are to each other, we
define a similarity metric based on the measured runtimes. For that we first
remove $84$ benchmarks that have not been solved by any solver. For the rest of
the benchmarks we use the PAR-2 score,
% based on the \num{5000}-second timeout
i.\,e., for each solver we set a score of \num{10000} for unsolved instances.
This leaves us with $316$ floats
$S_{1},\dots,S_{n}$ for each solver $S$. The \emph{similarity} of two solvers $A$
and $B$, normalized to the interval $[0,1]$ is defined
as:
% by Equation~\ref{eq-similarity}.

% \begin{equation}
% \label{eq-similarity}
\[
	similarity(A, B) =\quad 1 - \frac{\sum{|A_{i} - B_{i}|}}{316 \cdot \num{10000}}
\]
% \end{equation}
We calculated the \emph{similarity} between the $30$ solvers which achieved the
best PAR-2 score in the main track. The results are displayed in
Figure~\ref{fig-similarity-main} as a \emph{heat map}. Additionally the result
of hierarchically clustering the solvers based on their similarity is
illustrated as a \emph{dendrogram}. The height at which two solvers or clusters
are joined reflects how similar they are. For example enabling the \emph{trail
  saving} in \solver{CaDiCaL-alluip} has no impact on the runtime, resulting in
a similarity above \num{0.999}. Thus the two solvers are joined low in the
dendrogram.
A similar visualization was used to visualize the results of the SAT Race \cite{Balyo:2015:SATRace}.

One big cluster we can identify are the \solver{Maple}-decedents to the right,
all of which are modifications of the winners of the SAT Competition 2018 and
the SAT Race 2019.
With the exception of \solver{mergesat} and
maybe \solver{DurianSat} we observe a high similarity within the cluster.
% * 0.9645726159423101, 'exp_V_MLD_CBT_DL', 'Scavel' second highest similarity of all (not 2 configurations of the same solver)
% ** both based on MapleLCMDistChronoBT-dl-v2.2
% ** suggesting that the changes they made are either very similar or both do not have a high impact on the performance of the base solver in this test set.

% The two configurations of the \solver{Relaxed} by \emph{Zhang} and \emph{Cai}
use the same code base as a lot of the solvers in the Maple-Cluster but their
overall performance is better and closer to the CMS-Cluster.

The three CMS versions differ in their implementation of stochastic local search
(SLS) but overall have a similar performance.

The last cluster to the left of the heat map is comprised of the solvers
originally written by Biere. The three configurations of \solver{Kissat} form a
sub-cluster, even though the configurations are fairly distinct from each other.
% , specialized for satisfiable and unsatisfiable instances.
% One targets satisfiable instances, while the other is focused on unsatisfiable
% instances and a third alternating between them.

The other sub-cluster consists of modifications to the 2019 version of
\solver{CaDiCaL}. The new version CaDiCaL-2020 has a worse overall performance
than the solvers in this cluster and a comparatively high similarty to the
Maple-Cluster.

@@

The coefficient $\rho$ is in the interval $[-1, 1]$, where a rank correlation of
$1$ means that the two rankings are equal and $-1$ means that one ranking is the
reverse of the other. To give a better intuition for $\rho$, we list a few
modifications to the original ranking together with the resulting rank
correlation coefficient: The smallest change we can make is to switch the rank
of two adjacent solvers, resulting in a high rank correlation $\rho = 0.9999$.
Several small changes also result in a high rank correlation; repeating the same
modification as before $n/2$ times to switch every two adjacent solvers in rank
still gives a value of $\rho = 0.9974$. On the other hand, switching the highest
ranked solver (\solver{Kissat-sat}) with the lowest ranked solver, results in a
rank correlation of $\rho = 0.7602$. Moving \solver{Kissat-sat} to the bottom of
the ranking and moving every other solver up one rank gives a higher
$\rho = 0.8776$. Doing the same to all three \solver{Kissat} configurations
(with ranks 1, 2 and 11, now 46, 47 and 48) results in $\rho = 0.6839$.

The mean and standard deviation of the computed correlation coefficients are
depicted in Figure~\ref{fig:sampleAll}.

\begin{figure}[h]
  \label{fig:sampleAll}
  \includegraphics[width=\textwidth]{stability/ALL_random_smpling_correlations.png}
  \caption{Mean and standard deviation of rank correlation under removal of random instances, 50 samples for each size}
\end{figure}

The rank correlation is high even for relatively small samples.
The average rank correlation drops below $0.99$ only after randomly removing at least $95$ benchmarks, which is $30\%$ of the considered benchmark set. Accordingly, removing fewer benchmarks randomly has almost no effect on the ranking of the solvers.
Furthermore, removing fewer than $200$ ($63\%$) benchmarks, results in an average rank correlation above $0.96$

This suggest that the influence of the random selection in Algorithm~\ref{algo:select} on the solver ranking is limited.
% \ref{sec:byob}

The collected data cannot show, whether all of the benchmarks originally
submitted by the solver authors, have a systematic bias. Since each newly
submitted benchmark family originates from a different domain and is often the
result of current research, we can assume that the submitted families together
are representative. Figure~\ref{fig:sampleFamily} shows the rank correlation
coefficient resulting from removing a complete benchmark family instead of a
random subset. As expected removing a non random subset can have a higher impact
on the ranking even if it is small. Removing the 13 benchmarks from the
\emph{hgen}-family results in a (still high) rank correlation of $0.9895$. Additionally, the ranking of the top five solvers stays the same.

@@

>>>>

The Parallel track evaluates solver performance in wallclock time on multiple cores.
The benchmarks and the timeout (5000 seconds) are the same as in the Main track. 
Each solver run in this track was on a m4.16xlarge machine from Amazon Web Services.
These machines have 64 virtual cores and 256GB of memory. In contrast to the 
Main track, there is no proof logging for unsatisfiable benchmarks in the Parallel track.

@@

New this year was the Cloud track to evaluated distributed solvers. Similar to the
Parallel track, the solvers ran on the Amazon cloud, but using 100 machines of the
type m4.4xlarge, which have 16 virtual cores and 64GB of memory. 
Communication between the machines is possible using MPI and SSH.
The wallclock timeout is 1000 seconds and the same benchmarks were used as in
the Main and Parallel track.

@@

\subsection{Similarity of Solvers}
To investigate how similar the solvers of the Main track are to each other, we
define a similarity metric based on the measured runtimes. For that we first
remove $84$ benchmarks that have not been solved by any solver. For the rest of
the benchmarks we use the PAR-2 score,
% based on the \num{5000}-second timeout
i.\,e., for each solver we set a score of \num{10000} for unsolved instances.
This leaves us with $316$ floats
$S_{1},\dots,S_{n}$ for each solver $S$. The \emph{similarity} of two solvers $S$
and $S'$, normalized to the interval $[0,1]$ is defined
as:
% by Equation~\ref{eq-similarity}.

% \begin{equation}
% \label{eq-similarity}
\[
	similarity(S, S') =\quad 1 - \frac{\sum{|S_{i} - S'_{i}|}}{316 \cdot \num{10000}}
\]
% \end{equation}
We calculated the \emph{similarity} between the $30$ solvers which achieved the
best PAR-2 score in the main track. The results are displayed in
Figure~\ref{fig-similarity-main} as a \emph{heat map}. Additionally the result
of hierarchically clustering the solvers based on their similarity is
illustrated as a \emph{dendrogram}. The height at which two solvers or clusters
are joined reflects how similar they are. For example enabling the \emph{trail
  saving} in \solver{CaDiCaL-alluip} has no impact on the runtime, resulting in
a similarity above \num{0.999}. Thus the two solvers are joined low in the
dendrogram. A similar visualization was used for the results of the SAT Race 2015 \cite{Balyo:2015:SATRace}.

One big cluster we can identify are the \solver{Maple}-decedents to the right,
all of which are modifications of the winners of the SAT Competition 2018 and
the SAT Race 2019.
The similarity within the cluster is high, with the exception of
\solver{Scavel} and \solver{exp\_V\_MLD\_CBT\_DL}.
They form a subcluster
with a lower similarity to the rest, but a high similarity within. In fact the highest measured similarity,
 % of $0.96$
besides the aforementioned \solver{CaDiCaL-alluip}, is observed between them.
The two solvers are both based on
\solver{MapleLCMDistChronoBT-dl-v2.2} but have different authors.
The high similarity suggest that the
changes they made either result in very similar behavior or do not have a
significant impact on the runtime performance.

The two configurations of the \solver{Relaxed} by \emph{Zhang} and \emph{Cai}
use the same code base as a lot of the solvers in the Maple-Cluster but their
overall performance is better and closer to the CMS-Cluster.

Close to them are the three \solver{CMS} versions. They differ in their implementation of stochastic local search
(SLS) and have similar performance.
The 2020 version of \solver{CaDiCaL}
performed worse than the modifications based on the 2019 version and
does not quite fit into any cluster.

The last cluster to the left of the heat map is comprised of the other solvers
originally written by \emph{Biere}. Interesting to note is that the
\solver{Kissat} configuration specialized for unsatisfiable instances joins the
others in the cluster high in the dendrogram. In fact \solver{Kissat-unsat} has
the lowest average similarity to all other solvers in the top 30. This hints at
its importance for the use in an optimal portfolio.
% Average similarity
% | Kissat-unsat         |  21.75012381941168 |
% | Kissat               | 22.591060319234913 |
% | CaDiCaL-alluip       |  22.93588211785061 |
% | CaDiCaL-alluip-trail |  22.93631201417939 |
% | Kissat-sat           |  22.99738644487821 |

% The other sub-cluster consists of modifications to the 2019 version of
% \solver{CaDiCaL}.
% The new version CaDiCaL-2020 has a worse overall performance
% than the solvers in this cluster and a comparatively high similarty to the
% Maple-Cluster.

@@

The coefficient $\rho$ is in the interval $[-1, 1]$, where a rank correlation of
$1$ means that the two rankings are equal and $-1$ means that one ranking is the
reverse of the other. To give a better intuition for $\rho$, we list a few
modifications to the original ranking together with the resulting rank
correlation coefficient: The smallest change we can make is to switch the rank
of two adjacent solvers, resulting in a high rank correlation $\rho = 0.9999$.
Several small changes also result in a high rank correlation; repeating the same
modification as before $n/2$ times to switch all pairs of adjacent solvers in
rank still gives a value of $\rho = 0.9974$. On the other hand, switching the
highest ranked solver (\solver{Kissat-sat}) with the lowest ranked solver,
results in a rank correlation of $\rho = 0.7602$. Moving \solver{Kissat-sat} to
the bottom of the ranking and moving every other solver up one rank gives a
higher $\rho = 0.8776$. Doing the same to all three \solver{Kissat}
configurations (with ranks 1, 2 and 11, now 46, 47 and 48) results in
$\rho = 0.6839$.

The mean and standard deviation of the computed correlation coefficients are
depicted in Figure~\ref{fig:sampleAll}.
\begin{figure}[h]
  \label{fig:sampleAll}
  \includegraphics[width=\textwidth]{stability/ALL_random_smpling_correlations.png}
  \caption{Mean and standard deviation of rank correlation under removal of random instances, 50 samples for each size}
\end{figure}
The rank correlation is high even for relatively small samples.
The average rank correlation drops below $0.99$ only after randomly removing at least $95$ benchmarks, which is $30\%$ of the considered benchmark set. Accordingly, removing fewer benchmarks randomly has almost no effect on the ranking of the solvers.
Furthermore, removing fewer than $200$ ($63\%$) benchmarks, still results in an average rank correlation above $0.96$

This suggest that the influence of the random selection in Algorithm~\ref{algo:select} on the solver ranking is limited.
% \ref{sec:byob}
The collected data cannot show, whether all of the benchmarks originally
submitted by the solver authors, have a systematic bias. However since each
newly submitted benchmark family originates from a different domain and is often
the result of current research, we can assume that the submitted families
together are representative.

@@

Figure~\ref{fig:sampleFamily} shows the rank
correlation coefficient resulting from removing a complete benchmark family
instead of a random subset. As expected removing a non random subset can have a
higher impact on the ranking even if it is small. Removing the 13 benchmarks
from the \emph{hgen}-family results in a (still high) rank correlation of
$0.9895$. Additionally, the ranking of the top five solvers stays the same.
The individual removal of all other benchmarks families results in a rank correlation over $0.99$.
